---
title: "Statistical Inference Course Project-1"
author: "Chakra C"
date: "Sunday, May 24, 2015"
output: html_document
---

## OVERVIEW
In this project I will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. I will set lambda = 0.2 for all of the simulations. I will investigate the distribution of averages of 40 exponentials. Note that I will need to do a thousand simulations.

Simulate averages of 40 exponetials from 1000 random uniforms.

```{r}
set.seed(3)
lambda <- 0.2
simu_num <- 1000
average_size <- 40
sim <- matrix(rexp(simu_num*average_size, rate=lambda), simu_num, average_size)
row_means <- rowMeans(sim)
```

The distribution of sample means as shown below:

```{r echo=FALSE}
# Hist Graph for the histogram of averages
hist(row_means, breaks=50, prob=TRUE,
     main="Graph-1: Show the sample of avarage distributions drawn from 
     exponential distribution with Lambda = 0.2",
     xlab="")
# Averages of samples density
lines(density(row_means))
# theoretical center of distribution
abline(v=1/lambda, col="green")
# Averages of samples theoretical density 
xfit <- seq(min(row_means), max(row_means), length=100)
yfit <- dnorm(xfit, mean=1/lambda, sd=(1/lambda/sqrt(average_size)))
lines(xfit, yfit, pch=22, col="green", lty=2)
# Apply legend
legend('topright', c("Simulation", "Theoretical"), lty=c(1,2), col=c("black", "green"))
```

The sample means from distribution is centered at `r mean(row_means)`, theoretical center of the distribution is $\lambda^{-1}$ = `r 1/lambda`. sample means of variance is `r var(row_means)`, here theoretical variance
of the distribution is $\sigma^2 / n = 1/(\lambda^2 n) = 1/(0.04 \times 40)$ =
`r 1/(0.04 * 40)`. With this theoretical central limit is the average of samples normal distribution. From above Graph-1 explains the density calculated using the histogram, normal density plotted with theoretical mean and its variance values. 

Below Q-Q Graph-2 explains the normality.


```{r echo=FALSE}
qqnorm(row_means); qqline(row_means)
```

Examine coverage of the confidence interval for 
$1/\lambda = \bar{X} \pm 1.96 \frac{S}{\sqrt{n}}$

```{r echo=FALSE}
Lambda_values <- seq(4, 6, by=0.01)
Coverage <- sapply(Lambda_values, function(lamb) {
    mu_hats <- rowMeans(matrix(rexp(average_size*simu_num, rate=0.2),
                               simu_num, average_size))
    ll <- mu_hats - qnorm(0.975) * sqrt(1/lambda**2/average_size)
    ul <- mu_hats + qnorm(0.975) * sqrt(1/lambda**2/average_size)
    mean(ll < lamb & ul > lamb)
})

library(ggplot2)
qplot(Lambda_values, Coverage) + geom_hline(yintercept=0.95)
```

The rate parameter ($\lambda$) is estimated from 95% confidence intervals ($\hat{\lambda}$) are $\hat{\lambda}_{low} = \hat{\lambda}(1 - \frac{1.96}{\sqrt{n}})$ and 
$\hat{\lambda}_{upp} = \hat{\lambda}(1 + \frac{1.96}{\sqrt{n}})$.
From Graph-3 plot, selection for $\hat{\lambda}$ is around 5, sample mean avarage is within the confidence interval 95% of the time. And note that the real rate is $\lambda$ is 5.



